---
output:
  md_document:
    variant: markdown_github
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

```{r, echo=FALSE, results='asis'}
envreportutils::devex_badge("exploration")
```

[![Travis-CI Build Status](https://travis-ci.org/bcgov/rcaaqs.svg?branch=master)](https://travis-ci.org/bcgov/rcaaqs)

---

# rcaaqs

An [R](https://www.r-project.org/) package to faciliate the calculation of air quality 
metrics according to the Canadian Ambient Air Quality Standards
([CAAQS](http://www.ccme.ca/en/current_priorities/air/caaqs.html))

### Features

- General functions for doing things like formatting dates, filling in
  sequences, etc. 
- Functions for stepwise calculation of CAAQS metrics for
  different pollutants. Currently these are only complete for PM2.5 
  (annual and 24hr) metrics, and are in development for ozone. 
- Functions for assigning
  metrics for different pollutants into achievement and management categories.

### Installation

The package is not available on CRAN, but can be installed using
the [devtools](https://github.com/hadley/devtools) package:

```{r, eval=FALSE}
install.packages("devtools") # if not already installed

library(devtools)
install_github("bcgov/rcaaqs")

```

### Usage

This is a simple example using the included sample data set for PM<sub>2.5</sub>. For a fuller example with a larger dataset requiring more cleaning, see the [PM2.5 analysis we completed for 2011-2013](https://github.com/bcgov/pm25-caaqs-analysis).

```{r}
library(rcaaqs)
library(dplyr)

## Load sample data:
data("pm25_sample_data")
glimpse(pm25_sample_data)

## Calculate the data completeness criteria for the pm2.5 data:
pm_completeness <- pm_data_complete(pm25_sample_data, 
                                    by = c("ems_id", "site"))

## Columns are added showing the completeness evaluation annually 
## and quarterly
glimpse(pm_completeness)

## Compute daily averages by station (these are used for both the 24 hr 
## metric and the Annual metric:
avgdaily <- pm_daily_avg(pm25_sample_data, by = c("ems_id", "site"))
glimpse(avgdaily)

# Calculate the PM25 24 Hour Metric -----------------------------------

## Compute annual 98th percentile of 24h averages for each site
ann_98_per <- pm_98_percentile(avgdaily, by = "ems_id")
glimpse(ann_98_per)

## Then join the results to the data completeness data frame and select 
## only the results that are valid. This process could be more efficent
ann_98_per <- left_join(ann_98_per, pm_completeness, 
                        by = c("ems_id", "year"))
ann_98_per <- filter(ann_98_per, use_annual | (exceed & annual_valid))
glimpse(ann_98_per)

## Finally, calcualte the 24 hr CAAQS metric: 'The 3-year average of the 
## annual 98th percentile of the daily 24-hour average concentrations'
pm_caaq_daily <- pm_24h_caaq(ann_98_per, by = "ems_id", cyear = 2013)
glimpse(pm_caaq_daily)

# Calculate the PM25 Annual Metric ------------------------------------

## Compute the annaul average
annual_avg <- pm_annual_average(avgdaily, by = "ems_id")
glimpse(annual_avg)

## Then join the results to the data completeness data frame and select 
## only the results that are valid. This process could be more efficent
annual_avg <- left_join(annual_avg, pm_completeness, 
                        by = c("ems_id","year"))
annual_avg <- filter(annual_avg, use_annual)
glimpse(annual_avg)

## Finally, calculate the Annual CAAQS metric: The 3-year average of the
## annual average concentrations. 
pm_caaq_annual <- pm_annual_caaq(annual_avg, by = "ems_id", cyear = 2013)
glimpse(pm_caaq_annual)
```


### Project Status

The package is under active development. It is working well for the calculation
of PM<sub>2.5</sub> metrics, and the assignemnt of metrics into categories. We are working
on the ozone metric calculations.

### Getting Help or Reporting an Issue

To report bugs/issues/feature requests, please file an 
[issue](https://github.com/bcgov/rcaaqs/issues/).

### How to Contribute

If you would like to contribute to the package, please see our 
[CONTRIBUTING](CONTRIBUTING.md) guidelines.

Please note that this project is released with a [Contributor Code of Conduct](CODE_OF_CONDUCT.md). By participating in this project you agree to abide by its terms.

### License

    Copyright 2015 Province of British Columbia

    Licensed under the Apache License, Version 2.0 (the "License");
    you may not use this file except in compliance with the License.
    You may obtain a copy of the License at 

       http://www.apache.org/licenses/LICENSE-2.0

    Unless required by applicable law or agreed to in writing, software
    distributed under the License is distributed on an "AS IS" BASIS,
    WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
    See the License for the specific language governing permissions and
    limitations under the License.
    
This repository is maintained by [Environmental Reporting BC](http://www2.gov.bc.ca/gov/content?id=FF80E0B985F245CEA62808414D78C41B). Click [here](https://github.com/bcgov/EnvReportBC-RepoList) for a complete list of our repositories on GitHub.    
