---
output: github_document
html_preview: true
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

```{r, echo=FALSE, results='asis'}
envreportutils::devex_badge("exploration")
```

[![Travis-CI Build Status](https://travis-ci.org/bcgov/rcaaqs.svg?branch=master)](https://travis-ci.org/bcgov/rcaaqs)

---

# rcaaqs

An [R](https://www.r-project.org/) package to facilitate the calculation of air quality 
metrics according to the Canadian Ambient Air Quality Standards
([CAAQS](http://www.ccme.ca/en/current_priorities/air/caaqs.html))

### Features

- General functions for doing things like formatting dates, filling in
  sequences, etc. 
- Functions for stepwise calculation of CAAQS metrics&mdash;including implementing data completeness criteria&mdash;for
  different air pollutants. Currently these are complete for PM~2.5~ 
  (annual and 24hr), O~3~, NO~2~, and SO~2~ metrics. 
- Functions for assigning
  metrics for different pollutants into achievement and management categories.

### Installation

The package is not available on CRAN, but can be installed using
the [devtools](https://github.com/hadley/devtools) package:

```{r, eval=FALSE}
install.packages("devtools") # if not already installed

library(devtools)
install_github("bcgov/rcaaqs")

```

### Usage

This is a simple example using the included sample data set for PM<sub>2.5</sub>. 

```{r}
library(rcaaqs)
library(dplyr, warn.conflicts = FALSE)

## Load sample data:
data("pm25_sample_data")
glimpse(pm25_sample_data)

# Compute the daily average
avgdaily <- 
  pm_daily_avg(pm25_sample_data, 
               by = c("ems_id", "site"))
glimpse(avgdaily)

## Compute yearly 98 percentiles of daily average by station 
pm_98 <- 
  pm_yearly_98(avgdaily, 
               by = c("ems_id", "site"))
glimpse(pm_98)

## Compute annual averages by station 
annual_avg  <- 
  pm_yearly_avg(avgdaily, 
                by = c("ems_id", "site"))
glimpse(annual_avg)

## Calculate the Annual CAAQS metric: The 3-year average of the
## percentiles average concentrations. 
pm_98_just_valid <- pm_98[pm_98$exceed | pm_98$valid_year,]
pm_caaq_daily <- 
  pm_24h_caaq(pm_98_just_valid, by = "ems_id", cyear = 2013)
glimpse(pm_caaq_daily)

## Finally, calculate the Annual CAAQS metric: The 3-year average of the
## annual average concentrations. 
annual_avg_just_valid <- annual_avg[annual_avg$valid_year,]
pm_caaq_annual <- pm_annual_caaq(annual_avg_just_valid, by = "ems_id", cyear = 2013)
glimpse(pm_caaq_annual)
```

This is a simple example using the included sample data set for O<sub>3</sub>. 
```{r}
library(rcaaqs)
library(dplyr, warn.conflicts = FALSE)

## Load sample data:
data(o3_sample_data)
glimpse(o3_sample_data)

# Compute the daily rolling 8 hour average
rolling_avg <- o3_rolling_8hr_avg(o3_sample_data, by = c("ems_id", "site"))
glimpse(rolling_avg)

# Compute the daily maximum
daily_max_o3 <- o3_daily_max(rolling_avg, by = c("ems_id", "site"))
glimpse(daily_max_o3)

# Compute the 4th highest daily maximum
ann_4th_highest <- o3_ann_4th_highest(daily_max_o3, by = c("ems_id", "site"))
glimpse(ann_4th_highest)

# Compute the rolling three year average.
three_yr_avg <- o3_three_yr_avg(ann_4th_highest, by = c("ems_id", "site"))
glimpse(three_yr_avg)
```

This is a simple example using the included sample data set for SO<sub>2</sub>. 

```{r}
library(rcaaqs)
library(dplyr, warn.conflicts = FALSE)

## Load sample data:
data("so2_sample_data")
glimpse(so2_sample_data)

# Compute the daily average
so2_maxdaily <- 
  so2_daily_max(so2_sample_data, 
                by = c("ems_id", "site"))
glimpse(so2_maxdaily)

## Compute yearly 98 percentiles of daily average by station 
so2_99 <- 
  so2_yearly_99(so2_maxdaily, 
               by = c("ems_id", "site"))
glimpse(so2_99)

# Caclulate the three year statistics.
so2_three_yr <- 
  so2_three_yr_avg(so2_99,
                   by = c("ems_id", "site"))
glimpse(so2_three_yr)

# Compute the yearly average of all one-hour readings.
so2_yearly_avg <- 
  so2_avg_hourly_by_year(so2_sample_data,
                         by = c("ems_id", "site"))
glimpse(so2_yearly_avg)
```

This is a simple example using the included sample data set for NO<sub>2</sub>. 
```{r}
library(rcaaqs)
library(dplyr, warn.conflicts = FALSE)

## Load sample data:
data("no2_sample_data")
glimpse(no2_sample_data)

# Compute the daily average
no2_maxdaily <- 
  no2_daily_max(no2_sample_data, 
                by = c("ems_id", "site"))
glimpse(no2_maxdaily)

## Compute yearly 98 percentiles of daily average by station 
no2_98 <- 
  no2_yearly_98(no2_maxdaily, 
               by = c("ems_id", "site"))
glimpse(no2_98)

# Caclulate the three year statistics.
no2_three_yr <- 
  no2_three_yr_avg(no2_98,
                   by = c("ems_id", "site"))
glimpse(no2_three_yr)

# Compute the yearly average of all one-hour readings.
no2_yearly_avg <- 
  no2_avg_hourly_by_year(no2_sample_data,
                         by = c("ems_id", "site"))
glimpse(no2_yearly_avg)
```


You can view the CAAQS Achievement Levels as well as Management Levels in the included data frames. These are used internally in `pm_annual_caaq()` and `pm_24hr_caaq()` to assign metric values at each station to the appropriate CAAQS Achievement Level and Management Level:

```{r}
achievement_levels

management_levels
```


### Project Status

The package is under active development. 

### Getting Help or Reporting an Issue

To report bugs/issues/feature requests, please file an 
[issue](https://github.com/bcgov/rcaaqs/issues/).

### How to Contribute

If you would like to contribute to the package, please see our 
[CONTRIBUTING](CONTRIBUTING.md) guidelines.

Please note that this project is released with a [Contributor Code of Conduct](CODE_OF_CONDUCT.md). By participating in this project you agree to abide by its terms.

### License

    Copyright 2015 Province of British Columbia

    Licensed under the Apache License, Version 2.0 (the "License");
    you may not use this file except in compliance with the License.
    You may obtain a copy of the License at 

       http://www.apache.org/licenses/LICENSE-2.0

    Unless required by applicable law or agreed to in writing, software
    distributed under the License is distributed on an "AS IS" BASIS,
    WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
    See the License for the specific language governing permissions and
    limitations under the License.
    
This repository is maintained by [Environmental Reporting BC](http://www2.gov.bc.ca/gov/content?id=FF80E0B985F245CEA62808414D78C41B). Click [here](https://github.com/bcgov/EnvReportBC-RepoList) for a complete list of our repositories on GitHub.    
